Word Identification Under Multimodal Uncertainty

### 2017-02-07 Data_and_analysis/analysis_exp1.Rmd
new organization

---
title: "WordRec Experiment"
author: "A. Fourtassi & M. Frank"
date: "January 13, 2017"
output:
  html_document:
    number_sections: yes
    toc: yes
---

@@

Data, and exclusion

@@

d <- read_delim("data_experiment2_complete.txt", delim = " ") %>%
  filter(ID !="AR1IWBDA7MC86", ID !="A3GOKNZGZGHL45") %>% #took the experiment twice
  filter(type == "Task", problem=="No")  %>% #had a technical problem during the experiment
  filter(score > 0.5)  #got more less than 50% correct on the training trials

@@

joint <- d %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  multi_boot_standard(col = "answer")

ggplot(joint, 
       aes(x = sound_dist, y = mean, col = factor(concept_dist))) + 
  #geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                 # position = position_dodge(width = .1), size=0.3) + 
  geom_point(size=2) +
  geom_line(size=0.7) + 
  geom_line(data = sounds, aes(col = 1), col = "black",size=0.7) + 
  #geom_pointrange(data = sounds, aes(ymin = ci_lower, ymax = ci_upper, 
                             #        col = 1), col = "black", size=0.4)+
  geom_hline(data = concepts, aes(yintercept=mean, col = factor(concept_dist)), lty = 2)+
  xlab("Auditory distance") +ylab("% different")+
  scale_colour_discrete(name="Visual Dist")+
theme(legend.title = element_text(size=8))

ggplot(joint, 
       aes(x = concept_dist, y = mean, col = factor(sound_dist))) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1), size=0.3) + 
  geom_line(lty = 2) + 
  geom_line(data = concepts, aes(col = 1), col = "black") + 
  geom_pointrange(data = concepts, aes(ymin = ci_lower, ymax = ci_upper, 
                                     col = 1), col = "black", size=0.4) + 
  scale_colour_solarized() +
  xlab("Visual distance") +ylab("% different")+
  scale_colour_discrete(name="Auditory Dist")+
theme(legend.title = element_text(size=8))

@@

#Fit the uniomodal sound condition
sound_data <- d %>%
    filter(condition == "sound")

@@

#Fit the non-linear function

@@

##extract coefficient
eA_nl <- coef(sound_nl)["e"]
vrA_nl <- coef(sound_nl)["vrA"]

eV_nl <- coef(concept_nl)["e"]
vrV_nl <- coef(concept_nl)["vrV"]

eJ_nl <- coef(joint_nl)["e"]
vrJ_A_nl <- coef(joint_nl)["vrA_j"]
vrJ_V_nl <- coef(joint_nl)["vrV_j"]

@@

results <- boot(data= d, statistic = lmfit, R = 1)

@@

```
Modalitiy preference plot
```{r}
#I make a data frame 
model = c('baseline', 'fit')
mean=c(1.64,1.45)
ci_low=c(1.24,1.29)
ci_up=c(2.02,1.65)
pref=data.frame(mean, ci_low, ci_up,model)

ggplot(pref, 
       aes(x = model, y=mean)) +
geom_bar(stat = "identity", width=0.5, fill='grey')+
  geom_errorbar(aes(ymin = ci_low, ymax = ci_up), 
                  width = 0.1
                ,
                  position = position_dodge(width = 0.1))+
  theme(aspect.ratio = 1, axis.text=element_text(size=15))

```
Analysis of the unimodal cases.
```{r}

@@

#The fit
model_fit <- function (x,y) {
   1/(1 + (1-eJ_nl)*exp((-4/vrJ_A_nl)*x+(-4/vrJ_V_nl)*y+(8/vrJ_A_nl)+(8/vrJ_V_nl)))
}

#The baseline
model_base <- function (x,y) {
    1/(1 + (1-eJ_nl)*exp((-4/vrA_nl)*x+(-4/vrV_nl)*y+(8/vrA_nl)+(8/vrV_nl)))
}

@@

#Plot predictions of the baseline
mybase <- ms_all %>%
  filter(model=="base")

@@

ggplot(myfit, 
       aes(x = sound_dist, y = pred, col = factor(concept_dist))) + 
  geom_point(size=2.5) + 
  geom_line(size= 0.6) + 
  geom_line(data = sounds, aes(y=mean, col = 1), col = "black" ,size=0.8, lty=2) +
  geom_hline(data = concepts, aes(yintercept=mean, col = factor(concept_dist)), lty = 2)+
  #scale_colour_solarized() +
  xlab("Auditory distance") +ylab("% different")+
  scale_colour_discrete(name="Visual Dist")+
theme(legend.title = element_text(size=8))+
  scale_y_continuous(limits = c(0, 1))+
theme(aspect.ratio = 0.7)

@@

#Modality dominance coeficent
a_fit=vrJ_V_nl/vrJ_A_nl
a_base=vrV_nl/vrA_nl
a_Vdom=2*a_base
a_Sdom=0.5*a_base

>>>>

---
title: "WordRec Experiment 1"
author: "A. Fourtassi & M. C. Frank"
date: "January 13, 2017"
output:
  html_document:
    number_sections: yes
    toc: yes
---

@@

Data and exclusion

@@

#subject 32 and subject 59 are the same person
#subject 42 and subject 100 are the same person
#They will be excluded because they took the same experiment twice

d <- read_delim("data_exp1_anonym.txt", delim = " ") %>%
  filter(ID !="subject_32", ID !="subject_59", ID !="subject_42", ID !="subject_100") %>% 
  filter(type == "Task", problem=="No")  %>% #keep subjects who did not encounter a technical problem during the experiment
  filter(score > 0.5)  #.. and got more than 50% correct on the training trials

@@

joint <- d %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  multi_boot_standard(col = "answer")
```

Initial plot
```{r}
ggplot(joint, 
       aes(x = sound_dist, y = mean, col = factor(concept_dist))) + 
  #geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                 # position = position_dodge(width = .1), size=0.3) + 
  geom_point(size=2) +
  geom_line(size=0.7) + 
  geom_line(data = sounds, aes(col = 1), col = "black",size=0.7) + 
  #geom_pointrange(data = sounds, aes(ymin = ci_lower, ymax = ci_upper, 
                             #        col = 1), col = "black", size=0.4)+
  geom_hline(data = concepts, aes(yintercept=mean, col = factor(concept_dist)), lty = 2)+
  xlab("Auditory distance") +ylab("% different")+
  scale_colour_discrete(name="Visual Dist")+
theme(legend.title = element_text(size=8))

@@

sound_data <- d %>%
    filter(condition == "sound")

@@

#Fit the non-linear function
#

@@

##extract parameters

#The sound-ony bias term:
eA <- coef(sound_nl)["e"]
#The sound-only variance
vrA <- coef(sound_nl)["vrA"]

#The visual-ony bias term:
eV <- coef(concept_nl)["e"]
#The visual-ony variance:
vrV <- coef(concept_nl)["vrV"]

#The bimodal bias term:
eJ <- coef(joint_nl)["e"]

#The bimodal auditory variance:
vrJ_A <- coef(joint_nl)["vrA_j"]
#The bimodal visual variance:
vrJ_V <- coef(joint_nl)["vrV_j"]

@@

results <- boot(data= d, statistic = lmfit, R = 10000)

@@

Analysis of the unimodal cases.
```{r}

@@

#The bimodal fit
model_fit <- function (x,y) {
   1/(1 + (1-eJ)*exp((-4/vrJ_A)*x+(-4/vrJ_V)*y+(8/vrJ_A)+(8/vrJ_V)))
}

#The ideal observer model
model_base <- function (x,y) {
    1/(1 + (1-eJ)*exp((-4/vrA)*x+(-4/vrV)*y+(8/vrA)+(8/vrV)))
}

@@

#Plot predictions of the ideal observer model (or the baseline)
mybase <- ms_all %>%
  filter(model=="base")

@@

ggplot(myfit, 
       aes(x = sound_dist, y = pred, col = factor(concept_dist))) + 
  geom_point(size=2.5) + 
  geom_line(size= 0.6) + 
  geom_line(data = sounds, aes(y=mean, col = 1), col = "black" ,size=0.8, lty=2) +
  geom_hline(data = concepts, aes(yintercept=mean, col = factor(concept_dist)), lty = 2)+
  #scale_colour_solarized() +
  xlab("Auditory distance") +ylab("% different")+
  scale_colour_discrete(name="Visual Dist")+
theme(legend.title = element_text(size=8))+
  scale_y_continuous(limits = c(0, 1))+
theme(aspect.ratio = 0.7)
```

The correlations models vs. human data
```{r}

@@

#Modality dominance coeficent
a_fit=vrJ_V/vrJ_A
a_base=vrV/vrA
a_Vdom=2*a_base
a_Sdom=0.5*a_base

### 2017-02-07 Data_and_analysis/analysis_exp2.Rmd
new organization

---
title: "WordRec Experiment 1"
author: "A. Fourtassi & M. C. Frank"
date: "January 13, 2017"
output:
  html_document:
    number_sections: yes
    toc: yes
---

@@

Data and exclusion

#subject 32 and subject 59 are the same person
#subject 42 and subject 100 are the same person
#They will be excluded because they took the same experiment twice

d <- read_delim("data_exp1_anonym.txt", delim = " ") %>%
  filter(ID !="subject_32", ID !="subject_59", ID !="subject_42", ID !="subject_100") %>% 
  filter(type == "Task", problem=="No")  %>% #keep subjects who did not encounter a technical problem during the experiment
  filter(score > 0.5)  #.. and got more than 50% correct on the training trials

@@

Initial plot
```{r}
ggplot(joint, 
       aes(x = sound_dist, y = mean, col = factor(concept_dist))) + 
  #geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                 # position = position_dodge(width = .1), size=0.3) + 
  geom_point(size=2) +
  geom_line(size=0.7) + 
  geom_line(data = sounds, aes(col = 1), col = "black",size=0.7) + 
  #geom_pointrange(data = sounds, aes(ymin = ci_lower, ymax = ci_upper, 
                             #        col = 1), col = "black", size=0.4)+
  geom_hline(data = concepts, aes(yintercept=mean, col = factor(concept_dist)), lty = 2)+
  xlab("Auditory distance") +ylab("% different")+
  scale_colour_discrete(name="Visual Dist")+
theme(legend.title = element_text(size=8))

@@

#The bimodal auditory variance:
vrJ_A <- coef(joint_nl)["vrA_j"]
#The bimodal visual variance:
vrJ_V <- coef(joint_nl)["vrV_j"]

@@

ggplot(myfit, 
       aes(x = sound_dist, y = pred, col = factor(concept_dist))) + 
  geom_point(size=2.5) + 
  geom_line(size= 0.6) + 
  geom_line(data = sounds, aes(y=mean, col = 1), col = "black" ,size=0.8, lty=2) +
  geom_hline(data = concepts, aes(yintercept=mean, col = factor(concept_dist)), lty = 2)+
  #scale_colour_solarized() +
  xlab("Auditory distance") +ylab("% different")+
  scale_colour_discrete(name="Visual Dist")+
theme(legend.title = element_text(size=8))+
  scale_y_continuous(limits = c(0, 1))+
theme(aspect.ratio = 0.7)
```

>>>>

---
title: "WordRec Experiment 2"
author: "A. Fourtassi & M. C. Frank"
date: "January 13, 2017"
output:
  html_document:
    number_sections: yes
    toc: yes
---

@@

Data and exclusion
```{r}

#Skip the pre task trials 
#Filter subjects who had no technical problem
#Select subjects who got more than 50% correct answers on the pre task trials

d <- read_delim("data_exp2_anonym.txt", delim = " ") %>%
  filter(type == "Task", problem=="No") %>%
  filter(score > 0.5)
```

@@

Initial plot
```{r}

ggplot(joint, 
       aes(x = sound_dist, y = mean, col = factor(concept_dist))) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(width = .1), size=0.3) + 
  geom_line(lty = 2) + 
  geom_line(data = sounds, aes(col = 1), col = "black") + 
  geom_pointrange(data = sounds, aes(ymin = ci_lower, ymax = ci_upper, 
                                     col = 1), col = "black", size=0.4)+
  xlab("Auditory distance") +ylab("% different")+
  scale_colour_discrete(name="Visual Dist")+
theme(legend.title = element_text(size=8))

@@

#The bimodal auditory variance:
vrJ_A <- coef(joint_nl)["vrA_j"]
#The bimodal visual variance:
vrJ_V <- coef(joint_nl)["vrV_j"]
```

@@

ggplot(myfit, 
       aes(x = sound_dist, y = pred, col = factor(concept_dist))) + 
  geom_point(size=2.5) + 
  geom_line(size= 0.6) + 
  geom_line(data = sounds, aes(y=mean, col = 1), col = "black" ,size=0.8, lty=2) +
  geom_hline(data = concepts, aes(yintercept=mean, col = factor(concept_dist)), lty = 2)+
  #scale_colour_solarized() +
  xlab("Auditory distance") +ylab("% different")+
  scale_colour_discrete(name="Visual Dist")+
theme(legend.title = element_text(size=8))+
  scale_y_continuous(limits = c(0, 1))+
theme(aspect.ratio = 0.7)

### 2017-02-07 Writeup/main.tex
new organization

Identifying the visual referent of a spoken word -- that a particular insect is referred to by the word ``bee'' -- requires both the ability to process and integrate multi-modal input and the ability to reason under uncertainty. How do these tasks interact with one another? We introduce a task that allows us to examine how adults identify words under joint uncertainty in the auditory and visual modalities. We propose a ideal observer model of the task which provides an optimal baseline. Model predictions are tested in two experiments where word recognition is made under two kinds of uncertainty: category ambiguity and distorting noise. In both cases, the ideal observer model explains much of the variance in human judgments. But when one modality had noise added to it, human perceivers systematically preferred the unperturbed modality to a greater extent than the ideal observer model did.

@@

In the present work, we study, rather, the case of arbitrary mapping between sounds and visual objects. We test participants on their ability to process audio-visual stimuli and use them to recognize the underlying word. More precisely we study the case where both the word's form and the word's referent are ambiguous, and we examine the extent to which humans conform to, or deviate from the predictions of the ideal observer model. Moreover, some previous studies on audio-visual processing documented cases of modality preference, when people rely predominantly on the visual modality \cite<e.g., >{colavita1974} or the auditory modality \cite<e.g., >{sloutsky2003}. Thus, we will explore if participants in our task show evidence of a modality preference.

@@

This task is similar to the task introduced by \citeA{sloutsky2003} and used in subsequent research to probe audio-visual encoding. However, unlike this previous line of research, here participants are not asked whether the two audio-visual presentations are identical. Instead, the task is category-based: They are asked to press `same' if they think the second item (the test) belonged to the same category as the first (target) (e.g.,  dog-/aba/), even if there is a slight difference in the word, in the object, or in both. They are instructed to press `different' only if they think that the second stimulus was an instance of the other word category (cat-/ada/).

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.4\textwidth]{task1.png}
\caption{Overview of the task}
\label{fig:task}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=3.25in]{MyTask.png}
  \caption{Illustration of our model using simulated data. A word category is defined as the joint bivariate distribution of an auditory category (horizontal, bottom panel) and a visual semantic category (vertical, left panel). Upon the presentation of a word token $w$, participants guess whether it is sampled from the word category $W_1$ or from $W_2$. Decision threshold is where the guessing probability is 0.5.}
  \label{fig:space}
\end{figure}

@@

In sum, the posterior \ref{eq:pred} provides the predictions of ideal observer for how probabilities that characterize uncertainty in each modality can be combined to make categorical decision about bimodal input.

@@

For  auditory stimuli, we used the continuum introduced in \citeA{vroomen2004}, a 9-point /aba/--/ada/ speech continuum created by varying the frequency of the second (F2) formant in equal steps. We selected 5 equally spaced points from the original continuum by keeping the end-points (prototypes) 1 and 9, as well as points 3, 5, and 7 along the continuum. For visual stimuli, we used a morph continuum introduced in \citeA{freedman2001}: a continuous set of images was generated from three cat prototypes and three dog prototypes. From the original 14 points, we selected 5 points as follows: we kept the item that seemed most ambiguous (point 8), the 2 preceding points (i.e., 7 and 6) and the 2 following points (i.e., 9 and 10). The 6 and 10 points along the morph were quite distinguishable, and we took them to be our prototypes.

 % (Figure \ref{fig:morph})
%
% \begin{figure}[tp]
% \centering
% \includegraphics[width=0.4\textwidth]{morph.png}
% \caption{Visual stimuli for our experiments.}
% \label{fig:morph}
% \end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=3.25in]{unimodal.png}
\caption{Average human responses in the auditory-only condition (left), and visual-only condition (right). Error bars are 95\% confidence intervals. Solid lines represent unimodal logistic fits.}
\label{fig:unimodal}
\end{figure}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=.83\textwidth]{Exp.png}
\caption{Proportion of ``different'' judgments as a function of auditory distance. Solid lines represent average human responses (left), predictions of the ideal observer (middle), and the bimodal fit (left). Dashed lines represent average human responses in the unimodal conditions. Colors represent values in the visual continuum.}
\label{fig:Exp}
% \vspace{10ex}
\end{figure*}

% Let's take the auditory-only case as an example,
\subsubsection{Unimodal conditions} this is the case where the pictures were hidden, or where the sounds were muted.
Average categorization judgments and fits are shown in Figure \ref{fig:unimodal}. Auditory judgements were somewhat sharper and more pronounced than visual judgments. The fit was done as follows. For an ideal recognizer, the probability of choosing category 2 (that is, to answer ``different'') when presented with an audio instance $a$, is the posterior probability of this category $p(A_2|a)$. If we assume that both categories have equal variances, the posterior probability reduces to:

@@

\subsubsection{Bimodal condition}
We fit a model to human responses in the bimodal condition, collapsed across participants, finding $\epsilon=-0.32$, $\sigma^2_{Ab}=5.00$ and $\sigma^2_{Vb}=7.27$. The fit explained XX of the variance.

@@

We derived the predictions of the ideal observer model by using the values of the variances derived from the unimodal conditions, and the response bias derived from the bimodal condition, and by substituting these values into the expression of the posterior in Eq.~\ref{eq:pred}.
Figure \ref{fig:Exp} shows participants' responses in the bimodal condition (left), as well as the prediction of the ideal observer (middle), and the bimodal fit models (right).

@@

We used the same visual stimuli as in Experiment 1. We also used the same auditory stimuli, but we convolved each item with Brown noise of amplitude 1 using the audio editor Audacity (2.1.2).  The signal-to-noise ratio was XXX dB. The procedure was exactly the same as in the previous experiment, except that test stimuli were presented with the new noisy auditory stimuli.

@@

We fit a model to human responses in the bimodal condition, collapsed across participants. We estimated $\epsilon=-0.38$, $\sigma^2_{Vb}=5.21$, and  $\sigma^2_{Ab}+\sigma^2_{Nb}=9.84$.

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{preference.png}
\caption{Top: decision thresholds in the audio-visual space. Red dotted line is the prediction of the ideal observer. Blue dotted lines are cases where modality preference is twice as strong as the ideal observer. Solid line is the threshold derived from human data. Bottom: comparison of the threshold slope between the ideal observer and the fit to human data. Error bars are 95\% confidence intervals computed via non-parametric bootstrap.}
\label{fig:pref}
\end{figure}

@@

Overall, in both Experiments, the majority of the variance could be explained by an ideal observer that combined multi-modal information optimally. In the light of this main result, we can revisit some previous findings in the literature. For instance, \citeA{sloutsky2003} reported a dominance for the auditory modality in children. This dominance disappear or reverses in adults. Could this developmental difference be driven by changes in the intrinsic relative reliability of modalities across development? More work is needed to carefully examine this (and other) speculations, and more generally, to determine the extent to which the optimal combination account helps us better understand the mechanisms of word precessing and learning.

>>>>

Identifying the visual referent of a spoken word -- that a particular insect is referred to by the word ``bee'' -- requires both the ability to process and integrate multi-modal input and the ability to reason under uncertainty. How do these tasks interact with one another? We introduce a task that allows us to examine how adults identify words under joint uncertainty in the auditory and visual modalities. We propose an ideal observer model of the task which provides an optimal baseline. Model predictions are tested in two experiments where word recognition is made under two kinds of uncertainty: category ambiguity and distorting noise. In both cases, the ideal observer model explains much of the variance in human judgments. But when one modality had noise added to it, human perceivers systematically preferred the unperturbed modality to a greater extent than the ideal observer model did.

@@

In the present work, we study, rather, the case of arbitrary mapping between sounds and visual objects. We test participants on their ability to process audio-visual stimuli and use them to recognize the underlying word. More precisely we study the case where both the word's form and the word's referent are ambiguous, and we examine the extent to which humans conform to, or deviate from the predictions of the ideal observer model. Moreover, some previous studies on audio-visual processing documented cases of modality preference, when people rely predominantly on the visual modality \cite<e.g., >{colavita74} or the auditory modality \cite<e.g., >{sloutsky2003}. Thus, we will explore if participants in our task show evidence of a modality preference.

@@

This task is similar to the task introduced by \citeA{sloutsky2003} and used in subsequent research to probe audio-visual encoding. However, unlike this previous line of research, here participants are not asked whether the two audio-visual presentations are identical. Instead, the task is category-based: They are asked to press ``same'' if they think the second item (the test) belonged to the same category as the first (target) (e.g.,  dog-/aba/), even if there is a slight difference in the word, in the object, or in both. They are instructed to press ``different'' only if they think that the second stimulus was an instance of the other word category (cat-/ada/).

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.4\textwidth]{pictures/task.png}
\caption{Overview of the task}
\label{fig:task}
\end{figure}

@@

\begin{figure}[tp]
  \centering
  \includegraphics[width=3.25in]{pictures/model.png}
  \caption{Illustration of our model using simulated data. A word category is defined as the joint bivariate distribution of an auditory category (horizontal, bottom panel) and a visual semantic category (vertical, left panel). Upon the presentation of a word token $w$, participants guess whether it is sampled from the word category $W_1$ or from $W_2$. Decision threshold is where the guessing probability is 0.5.}
  \label{fig:space}
\end{figure}

@@

In sum, the posterior \ref{eq:pred} provides the ideal observer's predictions for how probabilities that characterize uncertainty in each modality can be combined to make categorical decision about bimodal input.

@@

For  auditory stimuli, we used the continuum introduced in \citeA{vroomen2004}, a 9-point /aba/--/ada/ speech continuum created by varying the frequency of the second (F2) formant in equal steps. We selected 5 equally spaced points from the original continuum by keeping the end-points (prototypes) 1 and 9, as well as points 3, 5, and 7 along the continuum. For visual stimuli, we used a morph continuum introduced in \citeA{freedman2001}: a continuous set of images was generated from three cat prototypes and three dog prototypes. From the original 14 points, we selected 5 points as follows: we kept the item that seemed most ambiguous (point 8), the 2 preceding points (i.e., 7 and 6) and the 2 following points (i.e., 9 and 10). The 6 and 10 points along the morph were quite distinguishable, and we took them to be our prototypes.
%(Figure \ref{fig:morph}).
%
%\begin{figure}[tp]
%\centering
%includegraphics[width=0.4\textwidth]{morph.png}
%\caption{Visual stimuli for our experiments.}
%label{fig:morph}
%\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=3.25in]{pictures/unimodal.pdf}
\caption{Average human responses in the auditory-only condition (left), and visual-only condition (right). A) represents data from Experiment 1, and B) data from Experiment 2. Error bars are 95\% confidence intervals. Solid lines represent unimodal logistic fits.}
\label{fig:unimodal}
\end{figure}

@@

\begin{figure*}[t]
\centering
\includegraphics[width=1\textwidth]{pictures/bimodal.pdf}
\caption{Proportion of ``different'' judgments as a function of auditory distance. Solid lines represent average human responses (left), predictions of the ideal observer (middle), and the bimodal fit (left). Dashed lines represent average human responses in the unimodal conditions. Colors represent values in the visual continuum. A) represents data from Experiment 1, and B) data from Experiment 2.}
\label{fig:Exp}
% \vspace{10ex}
\end{figure*}

% Let's take the auditory-only case as an example,
\subsubsection{Unimodal conditions} this is the case where the pictures were hidden, or where the sounds were muted.
Average categorization judgments and fits are shown in Figure (\ref{fig:unimodal}, A). Auditory judgements were somewhat sharper and more pronounced than visual judgments. The fit was done using the Nonlinear Least Squares (NLS) R package, as follows. For an ideal recognizer, the probability of choosing category 2 (that is, to answer ``different'') when presented with an audio instance $a$, is the posterior probability of this category $p(A_2|a)$. If we assume that both categories have equal variances, the posterior probability reduces to:

@@

\subsubsection{Bimodal condition}
We fit a model to human responses in the bimodal condition, collapsed across participants, finding $\epsilon=-0.32$, $\sigma^2_{Ab}=5.00$ and $\sigma^2_{Vb}=7.27$. The fit explained 94\% of total variance.

@@

We derived the predictions of the ideal observer model by using the values of the variances derived from the unimodal conditions, and the response bias derived from the bimodal condition, and by substituting these values into the expression of the posterior in Eq.~\ref{eq:pred}.
Figure (\ref{fig:Exp}, A) shows participants' responses in the bimodal condition (left), as well as the prediction of the ideal observer (middle), and the bimodal fit models (right).

@@

We used the same visual stimuli as in Experiment 1. We also used the same auditory stimuli, but we convolved each item with Brown noise of amplitude 1 using the audio editor Audacity (2.1.2).  The procedure was exactly the same as in the previous experiment, except that test stimuli were presented with the new noisy auditory stimuli.

@@

We fit a model to human responses in the bimodal condition, collapsed across participants. We estimated $\epsilon=-0.38$, $\sigma^2_{Vb}=5.21$, and  $\sigma^2_{Ab}+\sigma^2_{Nb}=9.84$. The fit explained 97\% of total variance.

@@

\begin{figure}[h]
\centering
\includegraphics[width=0.4\textwidth]{pictures/preference.png}
\caption{Top: decision thresholds in the audio-visual space. Red dotted line is the prediction of the ideal observer. Blue dotted lines are cases where modality preference is twice as strong as the ideal observer. Solid line is the threshold derived from human data. Bottom: comparison of the threshold slope between the ideal observer and the fit to human data. Error bars are 95\% confidence intervals computed via non-parametric bootstrap.}
\label{fig:pref}
\end{figure}

@@

Overall, in both Experiments, the majority of the variance could be explained by an ideal observer that combined multi-modal information optimally. In the light of this main result, we can revisit some previous findings in the literature. For instance, \citeA{sloutsky2003} reported a dominance for the auditory modality in children. This dominance disappears or reverses in adults. Could this developmental difference be driven by changes in the intrinsic relative reliability of modalities across development? More work is needed to carefully examine this (and other) speculations, and more generally, to determine the extent to which the optimal combination account helps us better understand the mechanisms of word precessing and learning.
